The role of advection versus diffusion in intraventricular transport

### 2025-02-04 paper/main.tex
Results 2, remove 'Computational flow modelling scenarios'

\subsection*{Computational flow modelling scenarios}

\mer{MER: @Halvor I suggest you move these two paragraphs into their respective
Results sections (Results 2 and 5).}
When simulating transport of solute in the ventricles,
we consider the same six different regions of interest that were
considered in the photoconversion experiments.
In region number one (the photoconversion region),
we use the function~\eqref{eq:photoconversion_curve} to enforce
the concentration such that it represents the experimental data
of photoconversion of proteins. We consider the three different diffusion
coefficients introduced in previous sections to simulate and compare results
for transport of molecules of varying sizes.

To simulate transport in the pathological case of region-specific cilia absence,
we consider three scenarios where we modify the boundary conditions of the CSF flow model.
The cilia tangential stress boundary condition for the boundary where we
remove cilia is changed to a free-slip boundary condition,
prescribing zero tangential forces to represent absence of cilia motion
in the region where we consider cilia removal.
We consider removal of cilia in the three separate cilia-covered regions
of the telencephalic (anterior) ventricle, the dorsal part of the diencephalic (middle) ventricle,
and the ventral part of the diencephalic ventricle~(\Cref{fig:figure1}d).
For the three scenarios, we simulate CSF flow modeled by the Stokes
equations~\eqref{eq:stokes_eqs} with the modified boundary conditions,
and use the resulting velocity field in transport simulations.

@@

\subsection*{Ventricular solute transport is advection-dominated for larger molecules}
We next ask what is the impact of these CSF flow patterns and
magnitudes on solute transport generally within the brain ventricles?
To address this question, we simulate the distribution and evolution
of different solutes subject to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(\Cref{fig:fig2}\fixme{e/x}). In particular, we consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to Exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
STM+GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s), and Dendra2 ($D_3 =
1.15 \times 10^{-10}$ m$^2$/s). As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the different-characteristics
solutes. For the smallest diffusion coefficient, representing
Exosomes, the advection induced by the CSF flow dominates diffusion
(\Cref{fig:fig2}a). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients, the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Cref{fig:fig2}\fixme{x-y}). Interestingly, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site (\Cref{fig:fig2}\fixme{z}).
\begin{figure}%[H]
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of Dendra ($D_1$), STM+GFP ($D_2$) and Exosomes ($D_3$).
    \mer{MER: I think you should show on the same color scale here in a) (the range of values lets you and then it is possible to compare). Split a into a, b, c. Show e) first (so that e) becomes a)}.
    \textbf{a.} Slices in the $xz$-plane ($y=0.140$ mm, the center of the geometry) showing the concentration field simulated with diffusion coefficients $D_1$, $D_2$ and $D_3$ for three time instants $t=45.0$ s, $t=112.6$ s, $t=450.5$ s.
    Note that the colorbars indicating the magnitude of $c$ have different limits for each column.
    \textbf{b--d.} Mean concentration $\cbar$ in each region of interest (ROI) as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively.
    \textbf{e.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{f.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

\mer{MER: Stopped here with the editing (except commenting out), assuming that Nathalie had something to say about the next paragraph.}

Transport of solute is quantified by the profile of the mean concentrations $\cbar_i(t)$ in the regions of interest (ROIs) $\Omega_i$ as functions of time~(\Cref{fig:fig2}b--d).
An illustration of the ROIs is provided in \Cref{fig:fig2}e. The dynamics of ROI 1 follow from imposing~\Cref{eq:photoconversion_curve}.
For ROIs 2--6, we observe more rapid concentration increases when the diffusion coefficient increases. 
For all three diffusion coefficients, the dynamics of ROI 6 is the slowest.
Region 5 exhibits a more rapid increase in mean concentration, and an even faster increase is observed for ROIs 2--4.
The profiles of the latter three regions are similar, displaying only small differences in all of the three plots.
\lyng{HH: Add some quantification in terms of percentages, e.g. final values of $\cbar_i$ for the different $D_i$?}
In region 6, the threshold is never exceeded for $D_1$, with the mean concentration close to zero at the end of the simulation. 
To quantify the balance of advection versus diffusion as transport mehcanisms on a global scale,
we calculate the Péclet numbers $\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the dominant transport mechanism for the molecules considered.

@@

\subsection*{Absence of ciliary motion affects local protein distribution}
To assess the impact of cilia motion on solute distribution,
we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant zebrafish (\emph{schmalhans} (\emph{smh})).
By plotting the means of $\Delta F$ (\Cref{fig:figure4_compare_control_mutant}a)
and the time to threshold
across the control and mutant cohorts (\Cref{fig:figure4_compare_control_mutant}b),
we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in $\sim X\%$ (ROI 2) and $\sim 45\%$ (ROI 4)
higher mean times-to-threshold in the \emph{smh}
mutants compared to controls (\Cref{fig:figure4_compare_control_mutant}b).

>>>>

@@

\subsection*{Ventricular solute transport is advection-dominated for larger molecules}
We next ask: what is the impact of these CSF flow patterns and
magnitudes on solute transport generally within the brain ventricles?
To address this question, we simulate the distribution and evolution
of different solutes subject to photoconversion within a region of
interest in the dorsal region of the middle ventricle (ROI 1)
(\Cref{fig:fig2}\fixme{e/x}). In particular, we consider the spread of
a solute concentration with a given molecular diffusion coefficient,
corresponding to Exosomes ($D_1 = 1.63 \times 10^{-12}$ m$^2$/s),
STM+GFP ($D_2 = 5.75 \times 10^{-11}$ m$^2$/s), and Dendra2 ($D_3 =
1.15 \times 10^{-10}$ m$^2$/s). As expected, transport occurs more
rapidly with increasing diffusion coefficient. Interestingly, the
distribution patterns differ between the \fixme{different-characteristics}
solutes. For the smallest diffusion coefficient, representing
Exosomes, the advection induced by the CSF flow dominates diffusion
(\Cref{fig:fig2}a). The vortex structures of the velocity field
carry over into the concentration field, indicating that transport
along the streamlines is more rapid than the diffusion across the
streamlines. The balance between diffusion and advection shifts with
increasing diffusion coefficients, the CSF flow structures become more
obfuscated, and the solute spreads more uniformly throughout the
ventricular geometry (\Cref{fig:fig2}\fixme{x-y}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to regions distal to the photoconversion site (\Cref{fig:fig2}\fixme{z}).
\begin{figure}%[H]
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}
    \caption{Simulated distribution and evolution after photoconversion of
    Dendra ($D_1$), STM+GFP ($D_2$) and Exosomes ($D_3$).
    \mer{MER: I think you should show on the same color scale here in a)
    (the range of values lets you and then it is possible to compare).
    Split a into a, b, c. Show e) first (so that e) becomes a)}.
    \textbf{a.} Slices in the $xz$-plane ($y=0.140$ mm,
    the center of the geometry) showing the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for three
    time instants $t=45.0$ s, $t=112.6$ s, $t=450.5$ s.
    Note that the colorbars indicating the magnitude of $c$ have different limits for each column.
    \textbf{b--d.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively.
    \textbf{e.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{f.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}
    \label{fig:fig2}
\end{figure}

To quantify the transport of solute, we plot the profile of the mean
concentrations $\cbar_i(t)$ as functions of time
(\cref{fig:figure2_compare_diffusion_coefficients}\fixme{b--d}) and
the times $\hat{t}$ when the mean concentrations $\cbar_i$ first exceed
a threshold value $\hat{c}$ (\cref{fig:figure2_compare_diffusion_coefficients}\fixme{f})
in 6 regions of interest (ROIs) (\cref{fig:figure2_compare_diffusion_coefficients}\fixme{e}).
The dynamics at the site of photoconversion (ROI 1)
is logarithmic and follows from imposing \eqref{eq:photoconversion_curve}.
For ROIs 2--6, we observe faster kinetics with higher diffusion coefficients
(\Cref{fig:figure2_compare_diffusion_coefficients}\fixme{b--d}),
which results in smaller time to threshold
(\Cref{fig:figure2_compare_diffusion_coefficients}\fixme{f}).
For all three diffusion coefficients, the dynamics of ROI 5
is significantly delayed compared to ROI 2--4,
and ROI 6 is the slowest. Notably, for $D_1$ the mean concentration
in ROI 6 is close to zero at the end of simulation.
To quantify the balance of advection versus diffusion as transport mehcanisms,
we calculate the Péclet numbers $\mathrm{Pe}_1=519$,
$\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.
Since $\mathrm{Pe}>1$, this indicates that advection is the
dominant transport mechanism on a \fixme{global} scale.

@@

\subsection*{Absence of ciliary motion affects local protein distribution}
Since CSF flow is highly disturbed in the absence of motile cilia, we
want to assess the impact of cilia motion on solute distribution. To this end,
we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant zebrafish (\emph{schmalhans} (\emph{smh})).
By plotting the means of $\Delta F$ (\Cref{fig:figure4_compare_control_mutant}a)
and the time to threshold
across the control and mutant cohorts (\Cref{fig:figure4_compare_control_mutant}b),
we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in $\sim X\%$ (ROI 2) and $\sim 45\%$ (ROI 4)
higher mean times-to-threshold in the \emph{smh}
mutants compared to controls (\Cref{fig:figure4_compare_control_mutant}b).
